# GIS on StackExchange

Here we will review basic CSV reading and writing to get comfortable with R and R Studio while looking at a dataset on the top questions on GIS StackExchange. 

If you're already familiar with base R, and dplyr, you can skip this chapter. 

## Analysis Goals

For questions on GIS Stackechange, summarize the relationship between:
- View Count, 
- Scores, 
- top questions

## Learning Goals

- Read a CSV from a URL
- how to use R to filter and visualize data. 
- the two kinds of R syntax that are common 

## CSV Import

Lets import the top 5000 most viewed questions on GIS StackExchange. 

```{r, cache=TRUE}
library(readr)
se_top_gis <- read_csv("http://data.stackexchange.com/gis/csv/1052797")
```

## R Grammar

You'll see two kinds of R grammar (syntax) in the wild. 

Sometimes they will be mixed. 

Some examples follow. 

### Base R syntax

Print the names, the top of the data set (filtered to the columns Title and Viewcount), and a histogram of the Viewcount, in 100 bins. 

```{r, cache=TRUE}
names(se_top_gis)
head(se_top_gis[,c('Title','ViewCount')])
hist(se_top_gis$ViewCount, n=100)
```

### dplyr syntax

Here we'll do the same thing with dplyr. 

dplyr's authors describe it as: "an opinionated grammar of data manipulation."

I think it reads like SQL in reverse, with unix pipes. 

R Studio has great cheat sheets [here](https://www.rstudio.com/resources/cheatsheets/).

```{r, results=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
```

```{r, cache=TRUE}
se_top_gis %>%
  select(Title,ViewCount, Score,CreationDate) %>%
    arrange(desc(ViewCount)) %>%
      head() %>%
        knitr::kable()

ggplot(se_top_gis, aes(ViewCount)) +
  geom_histogram()
```

Note how we've called out the `knitr` package before `kable`. You'll also see this in the wild with R scripts. This shorthand allows us to directly call a function from a package without loading the library. Doing this can clarify your script if you find yourself loading a lot of libraries at the beginning, from which you only use 1 function.   

## Visualization

Can we say something about how the number of views that a question has will relate to its score? What kind of relationship would we expect between the two variables?

```{r, cache=TRUE}
ggplot(se_top_gis) +
  geom_point(mapping = 
               aes(x = Score, 
                   y = ViewCount))
```

Lets see how the creationdate relates to the number of views. 

```{r, cache=TRUE}
plot1 <- ggplot(se_top_gis) +
  geom_point(mapping = 
               aes(x = Score, 
                   y = log(ViewCount), 
                   color=CreationDate))

plot1
```

```{r, cache=TRUE}
plot1 + facet_grid(lubridate::year(CreationDate) ~ .)
```

### Exercises 

#### 1

Make the histogram from ggplot above better

#### 2 

Edit the following to list the top 20 questions by Score and ViewCount. 

```{r, cache=TRUE}
se_top_gis %>%
  arrange(desc(Score,ViewCount)) %>% 
      filter(between(row_number(), 1, 10)) %>%
        knitr::kable()
```

#### 3

Below is variable with a url pointing to the top 50 queries on stats.stackechange by views.

Try reading the data and listing the top question titles and the number of views.  

```{r, cache=TRUE}
stats_top_50_url <- "https://gist.githubusercontent.com/tbuckl/9fd7e14e14f127536666b814991bbe0d/raw/0b629370d1cd2bc9948f54768fa268688f588e39/top_50_stats_questions_by_views.csv"
```



